### 2014-02-11 msc-thesis/chapters/1/chapter-intro.tex
MSC Added changes to thesis.

%Describe difference between direct and indirect
Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
which influences behavior negatively in another location in the system. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
previous interviews and surveys conducted with software developers have shown a pattern that developers 
of a software project view indirect conflict awareness  as a high priority issue in their 
development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
is required to address this developer concern.

@@

% Issues with IC tools of information overload etc.
While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, tools eventually being
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
also compounded by the scale of many software development projects leading to further information overload.
Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
come up.

% There are 4 studies
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. In order to fully understand indirect conflicts, this dissertation is composed
of 4 studies of which explore different components of indirect conflicts.

@@

% Study 2 introduction
Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool.
In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationship 
through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to full understand 
the causes of these indirect conflict tool issues, a third study was conducted.

@@

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I also set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Through this exploration, I looked to find what can be accomplished moving forward with indirect conflicts
in both research and industry. I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
a fourth and final study.

% Study 4 introduction
In order to begin to address the gap in software evolution analytical tools discovered in study 3, I turn my analysis to the notions 
of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts (as will be seen in Study 3), this study investigates the question:
``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
stability measure?}''. I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied.

>>>>

%Describe difference between direct and indirect
Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influence negatively. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
previous interviews and surveys conducted with software developers have shown a pattern that developers 
of a software project view indirect conflict awareness  as a high priority issue in their 
development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
is required to address this developer concern.

@@

% Issues with IC tools of information overload etc.
While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, tools eventually being
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
also compounded by the scale of many software development projects leading to further information overload.
Lastly, social factors such as Cataldo et al's.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
come up.

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
in the are of indirect conflicts. To accomplish this goal, I have researcher the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken in academia to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting the following four studies.

\section{Research Methodology}

In order to address the research goal as laid out in the previous section, I have conducted 4 studies which will
now be briefly outlined. Each study builds off the previous one and has research questions informed from the
findings of the previous study.

@@

% Study 2 introduction
Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool while
answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
developers who's changing technical dependencies statistically relate to bugs?}''.
In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationship 
through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to fully understand 
the causes of these indirect conflict tool issues, a third study was conducted.

@@

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I also set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
in both research and industry. This study asked the following 3 research questions: 
\textit{What events or conditions lead to indirect conflicts?}
\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
\textit{What do developers want from future indirect conflict tools?}

I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. The study findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
a fourth and final study.

% Study 4 introduction
In order to begin to address the gap in software evolution analytical tools discovered in study 3, I turn my analysis to the notions 
of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts (as will be seen in Study 3), this study investigates the question:
``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
stability measure?}''. I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied.

### 2014-02-11 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC Added all 4 studies to the thesis.
MSC intro now has big chunks in it.
MSC Added changes to thesis.

% Research question
With the power of technical dependencies in predicting software failures, the question I
investigated in this study was : 
``\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
statistically relate to bugs?}''

% Explain our brief intended approach
This study explains the approach used to locate these pairs of developers in developer networks.
The process utilizes code changes and the call hierarchies effected  to find patterns of developer 
relationships in successful and failed code changes. As it will be seen, we found 27 statistically
significant failure inducing developer pairs. These developer relationships can also be used
to promote the idea of leveraging socio-technical congruence, a measure of coordination compared
to technical dependencies amongst stakeholders, to provide coordination recommendations.

@@

% Technical network figure
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{figures/TecNetwork}
\caption{A technical network for a code change. Carl has changed method getX() which is being
called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
\end{figure}

>>>>

% Research question
With the power of technical dependencies in predicting software failures, the question I
investigated in this study was:

\begin{description}
        \item[RQ\namedlabel{itm:rq1-pairs}{RQ1-Pairs}] \textit{Is it possible to identify pairs 
        of developers whose technical dependencies in code changes statistically relate to bugs?}
\end{description}

% Explain our brief intended approach
This study explains the approach used to locate these pairs of developers in developer networks.
The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. As it will be seen, I found 27 statistically
significant failure inducing developer pairs. These developer relationships can also be used
to promote the idea of leveraging socio-technical congruence, a measure of coordination compared
to technical dependencies amongst stakeholders, to provide coordination recommendations.

@@

% Technical network figure
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{figures/TecNetwork}
\caption{A technical network for a code change. Carl has changed method getX() which is being
called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
\end{figure}

### 2014-02-11 msc-thesis/chapters/2/sec-impact.tex
MSC Added changes to thesis.

%Outline of study
In this study, I report on research into supporting developer pairs in indirect conflicts
and present the design, implementation, and future evaluation of the tool \textit{Impact},
a web based tool that aims at detecting indirect conflicts among developers
and notifying the appropriate members involved in these conflicts.
By leveraging technical relationships inherent of 
software projects with method call graphs~\cite{Lakhotia:1993:CCM}
as well as detecting changes
to these technical relationship through software configuration management
(SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
alert developers involved in such conflicts in task awareness while limiting information
overload by using design by contract~\cite{Meyer:1988} solutions to method design. While this study
outlines \textit{Impact's} specific implementation, its design is rather
generic and can be implemented in similar indirect conflict awareness tools.

\section{Related Work}
Although there is an abundance of awareness tools developed in research
today, only a handful have made an attempt to examine indirect conflicts.
Here, I will outline four of the forefront projects in indirect conflicts
and how these projects have influenced the decision making process in
the design and implementation of \textit{Impact}.

@@

>>>>

%Outline of study
In this study, I report on research into supporting developer pairs in indirect conflicts
and present the design, implementation, and future evaluation of the tool \textit{Impact},
a web based tool that aims at detecting indirect conflicts among developers
and notifying the appropriate members involved in these conflicts. Through Impact and
its evaluation I ask:

\begin{description}
        \item[RQ\namedlabel{itm:rq1-impact}{RQ1-Impact}] \textit{Can indirect conflicts be supported 
        through an awareness mechanism which leverages pairs of developers who's changing technical 
        dependencies statistically relate to bugs?}
\end{description}

By leveraging technical relationships inherent of 
software projects with method call graphs~\cite{Lakhotia:1993:CCM}
as well as detecting changes
to these technical relationship through software configuration management
(SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
alert developers involved in such conflicts in task awareness while limiting information
overload by using design by contract~\cite{Meyer:1988} solutions to method design. While this study
outlines \textit{Impact's} specific implementation, its design is rather
generic and can be implemented in similar indirect conflict awareness tools.

After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

\subsection{Related Work}
Although there is an abundance of awareness tools developed in research
today, only a handful have made an attempt to examine indirect conflicts.
Here, I will outline four of the forefront projects in indirect conflicts
and how these projects have influenced the decision making process in
the design and implementation of \textit{Impact}.

@@

\subsection{Threats to Validity}
% TODO

### 2014-02-11 msc-thesis/chapters/3/chapter-exploration.tex
MSC Added changes to thesis.

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

@@

Social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence, have also been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, false positives, dependencies (in developer organizational structure), and scalability 
come up.

While these issues of information overload, false positives, dependencies, and scalability continue to come up
in most indirect conflict tools, only a handful of attempts have been made at rectifying these issues or
finding the root causes~\cite{Holmes:2010:CAR,Kim:2011:ESA}. 
In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, I will proceed by taking a step back and determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
causes of source code changes in indirect conflicts, we may be able to create indirect conflict tools which have filtered
monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
I then set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
the same time are not using the tools put forth by academia, I wish to find what they use to mitigate indirect
conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
the hopes of a higher adoption rate of tools produced by academia.
Finally, I look to find what can be accomplished moving forward with indirect conflicts
in both research and industry.

I restate the research goals of this thesis for ease of the reader:

\begin{description}
  \item \textit{What events or conditions lead to indirect conflicts?}
  \item \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
  \item \textit{What do developers want from future indirect conflict tools?}
\end{description}

To answer these research questions, I performed a study (Section~\ref{study:exploration}) in which I interviewed 19 developers from across 12 commercial and open source projects, 
followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews. Based on the findings (to be seen) to the aforementioned research questions, 
I also performed a secondary complimentary study of software evolutionary trends (Section~\ref{study:apie}) to solidify
developer opinion and provide a starting point for the future development of indirect conflict tools.

>>>>

@@

Social factors such as Cataldo et al's.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence, have also been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, false positives, dependencies (in developer organizational structure), and scalability 
come up.

While these issues of information overload, false positives, dependencies, and scalability continue to come up
in most indirect conflict tools, only a handful of attempts have been made at rectifying these issues or
finding the root causes~\cite{Holmes:2010:CAR,Kim:2011:ESA}. 
In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts in this study, I examine and determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
causes of source code changes in indirect conflicts, we may be able to create indirect conflict tools which have filtered
monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
I then determine what mitigation strategies developers currently use as opposed to those created
by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
the same time are not using the tools put forth by academia, I wish to find what they use to mitigate indirect
conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
the hopes of a higher adoption rate of tools produced by academia.
Finally, I examine what can be accomplished moving forward with indirect conflicts
in both research and industry.

To explore and answer the research goals listed above, I performed a study (Section~\ref{study:exploration}) in which I 
interviewed 19 developers from across 12 commercial and open source projects, 
followed by a confirmatory questionnaire of 78 developers, and 5 confirmatory interviews.

Based on some of the findings (to be seen in details in Section~\ref{study:exploration}) I performed a follow up study
which did not relate directly to the themes of this dissertation but helped strengthen the results found in Study 3.
Some results of Study 3 showed that indirect conflict tools should take into account a contextual setting of development
progression in software projects to better inform developers of potential indirect conflicts. In other words, an indirect
conflict tool should be able to tell what phase of development inside the development life cycle a project is currently
active in. In order to better explore and support this finding, I performed a complimentary study of software evolutionary trends (Section~\ref{study:apie}).
In Study 4, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.

### 2014-02-11 msc-thesis/chapters/3/sec-apie.tex
MSC Added changes to thesis.

Release points are a vital milestone of software projects. From major releases of a Waterfall based project to minor iterations of an Agile development,
releases form an interesting single point of a project's development history.
Third party users (outside developers) of a system often only see a product at a release point either major or minor,
and expect the system to come with a sense of reliability and stability at this point. Developers often expect to be able to upgrade a library to
a newer version without having to make major revisions to their own projects to accommodate the upgrade (unless of course patch notes detailing major
changes are released). However, major and minor releases of a library or software resource can cause software quality to degrade in
a third party application as indirect conflicts may occur. The knowledge as to when a project is ready for public usage
as to its reliability, quality, stability, and thus indirect conflict proneness can be a difficult decision to make for most project owners or maintainers.

@@

I turn my analysis to the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts, this study investigates the question:
``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
stability measure?}''.

In this study, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near and major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.

\section{Related Work}
\label{sec:rel}
While very little has been published about release quality studies and stability (especially in regards to
indirect conflicts), there have been a few studies which attempt to address the issues directly
or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a project's behavior
around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be performed.
Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
at major and minor releases, showing how test and production code can evolve at different rates and times. These results show that production
and test code should be handled as different cases for a stability measure around major releases.

@@

I took ChangeDistiller and applied it across 10 open source Java projects. Java projects were necessary because ChangeDistiller only works for Java source code.
For each of the projects, I obtained the software configuration management (SCM) system
which is used to store all source code changes of a project. When it was necessary, I converted some forms of SCM system to Git in order to reduce implementation
burdens of using multiple SCMs. Once the SCM was obtained, I used ChangeDistiller and iterated over every commit found in a project's git master branch. I stored
34 of ChangeDistiller's built in source code change types for each commit. I noted how many of each change type was performed in each commit and stored that information
in a PostgreSQL database. In order to filter and protect my results, I manually inspected the 10 Java projects studied in order to identify code built for test
purposes.
I separated changes to this test code from all other code to ensure my results only focused on real implementation while allowing us to study changes to
test based code separately.

@@

To answer my research question, I conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.

@@

>>>>

Release points are a vital milestone of software projects. From major releases of a Waterfall based project to iterations of an Agile development,
releases form an interesting single point of a project's development history.
Third party users (outside developers) of a system often only see a product at a release point either major or minor,
and expect the system to come with a sense of reliability and stability at this point. Developers often expect to be able to upgrade a library to
a newer version without having to make major revisions to their own projects to accommodate the upgrade (unless of course patch notes detailing major
changes are released). However, major and minor releases of a library or software resource can cause software quality to degrade in
a third party application as indirect conflicts may occur. The knowledge as to when a project is ready for public usage
as to its reliability, quality, stability, and thus indirect conflict proneness can be a difficult decision to make for most project owners or maintainers.

@@

In this study, I examine the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts, this study investigates the question:

\begin{description}
        \item[RQ\namedlabel{itm:rq1-apie}{RQ1-APIE}] \textit{What trends exist in source code changes surrounding major releases of open 
        source projects as a notion towards a project stability measure?}
\end{description}

In this study, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.

\subsection{Related Work}
\label{sec:rel}
While very little has been published about release quality studies and stability (especially in regards to
indirect conflicts), there have been a few studies which attempt to address the issues directly
or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a project's behavior
around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be performed.
Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
at major and minor releases, showing how test and production code can evolve at different rates and times. These results show that production
and test code should be handled as different cases for a stability measure around major releases.

@@

I conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
because ChangeDistiller only works for Java source code and because of their high use amongst other Java projects and to study specific ecosystems 
of projects and their evolution trends.

For each of the projects, I obtained the software configuration management (SCM) system
which is used to store all source code changes of a project. When it was necessary, I converted some forms of SCM system to Git in order to reduce implementation
burdens of using multiple SCMs. Once the SCM was obtained, I used ChangeDistiller and iterated over every commit found in a project's git master branch. I stored
34 of ChangeDistiller's built in source code change types for each commit. I noted how many of each change type was performed in each commit and stored that information
in a PostgreSQL database. In order to filter and protect my results, I manually inspected the 10 Java projects studied in order to identify code built for test
purposes.
I separated changes to this test code from all other code to ensure my results only focused on real implementation while allowing us to study changes to
test based code separately.

@@

@@

I found 9 major change trends which surround major releases in the open source
projects studied. 4 change trends found that occur before major releases are: added private classes, 
changed test method signatures, changed documentation, and removed test classes.
5 change trends found that occur after major release are: added test methods, changed test classes, removed public methods, removed
private classes, and removed private methods.
